---
title: "Proyect_Bruno_Tarea_1"
author: "Enzo Aravena"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Reading the data

```{r}
library(tidyverse)
library(readr)
```


```{r}
ArchivoC_Adm2020 <- read_csv("Data/ArchivoC_Adm2020.csv")
ArchivoC_Adm2019 <- read_csv("Data/ArchivoC_Adm2019.csv")
ArchivoC_Adm2018 <- read_csv("Data/ArchivoC_Adm2018.csv")
ArchivoC_Adm2017 <- read_csv("Data/ArchivoC_Adm2017.csv")
ArchivoC_Adm2016 <- read_csv("Data/ArchivoC_Adm2016.csv")
ArchivoC_Adm2015 <- read_csv("Data/ArchivoC_Adm2015.csv")
ArchivoC_Adm2014 <- read_csv("Data/ArchivoC_Adm2014.csv")
ArchivoC_Adm2013 <- read_csv("Data/ArchivoC_Adm2013.csv")
ArchivoC_Adm2012 <- read_csv("Data/ArchivoC_Adm2012.csv")

```

## Delimiter

```{r}


Adm2020 <- ArchivoC_Adm2020 <- read_delim("Data/ArchivoC_Adm2020.csv", 
                               delim = ";")
Adm2019 <- ArchivoC_Adm2019 <- read_delim("Data/ArchivoC_Adm2019.csv", 
                               delim = ";")
Adm2018 <- ArchivoC_Adm2018 <- read_delim("Data/ArchivoC_Adm2018.csv", 
                               delim = ";")
Adm2017 <- ArchivoC_Adm2017 <- read_delim("Data/ArchivoC_Adm2017.csv", 
                               delim = ";")
Adm2016 <- ArchivoC_Adm2016 <- read_delim("Data/ArchivoC_Adm2016.csv", 
                               delim = ";")
Adm2015 <- ArchivoC_Adm2015 <- read_delim("Data/ArchivoC_Adm2015.csv", 
                               delim = ";")
Adm2014 <- ArchivoC_Adm2014 <- read_delim("Data/ArchivoC_Adm2014.csv", delim = ";")
Adm2013 <- ArchivoC_Adm2013 <- read_delim("Data/ArchivoC_Adm2013.csv", delim = ";")
Adm2012 <- ArchivoC_Adm2012 <- read_delim("Data/ArchivoC_Adm2012.csv", delim = ";")
```

```{r}
Adm2013 <- ArchivoC_Adm2013 <- read_delim("Data/ArchivoC_Adm2013.csv", delim = ";")
Adm2012 <- ArchivoC_Adm2012 <- read_delim("Data/ArchivoC_Adm2012.csv", delim = ";")
```



## 2020 (where 0 are treated as NA)

```{r}
Adm2020_na <- Adm2020 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  )
```


```{r}
glimpse(Adm2020_na)

Adm2020_na %>%
  group_by(RBD) %>%
  summarise(count_LENG = sum(!is.na(LENG_ACTUAL)))
```


```{r}
Adm2020_na <- Adm2020_na %>% 
  filter(SITUACION_EGRESO == 1)
Adm2020_na <- Adm2020_na %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2020_na_grouped <- Adm2020_na %>%
  group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
    )
Adm2020_na_grouped <- Adm2020_na_grouped %>%
  left_join(Adm2020_na %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")


```


## 2020 (where 0 are counted in the calculations)

```{r}
Adm2020 <- Adm2020 %>% 
  filter(SITUACION_EGRESO == 1)

Adm2020 <- Adm2020 %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)
# I want to compute the median per rbd in the columns LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM. I will use the group_by and summarise functions from the dplyr package to do this. I will group by RBD and then compute the mean for each of the columns mentioned above. I will also add a new column that counts the number of students in each RBD group. I also want to compute the percentile 10, 25, 75, and 90 for each of the columns mentioned above. I will use the quantile function from the base R package to do this. I will add new columns for each of the percentiles computed. I will name the columns as follows: LENG_ACTUAL_p10, LENG_ACTUAL_p25, LENG_ACTUAL_MEDIAN, LENG_ACTUAL_p75, LENG_ACTUAL_p90, MATE_ACTUAL_p10, MATE_ACTUAL_p25, MATE_ACTUAL_MEDIAN, MATE_ACTUAL_p75, MATE_ACTUAL_p90, HCSO_ACTUAL_p10, HCSO_ACTUAL_p25, HCSO_ACTUAL_MEDIAN, HCSO_ACTUAL_p75, HCSO_ACTUAL_p90, CIEN_ACTUAL_p10, CIEN_ACTUAL_p25, CIEN_ACTUAL_MEDIAN, CIEN_ACTUAL_p75, CIEN_ACTUAL_p90, PTJE_NEM_p10, PTJE_NEM_p25, PTJE_NEM_MEDIAN, PTJE_NEM_p75, PTJE_NEM_p90.

Adm2020_grouped <- Adm2020 %>% 
  group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    ) 

Adm2020_grouped <- Adm2020_grouped %>% 
  left_join(Adm2020 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")

```

## 2019 (Where 0 are treated as NA)

```{r}
Adm2019_na <- Adm2019 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  )

Adm2019_na <- Adm2019_na %>%
  filter(SITUACION_EGRESO == 1)
Adm2019_na <- Adm2019_na %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2019_na_grouped <- Adm2019_na %>%
  group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
    )
Adm2019_na_grouped <- Adm2019_na_grouped %>%
  left_join(Adm2019_na %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")


```



## 2019 (where 0 are counted in the calculations)

```{r}
Adm2019 <- Adm2019 %>% 
  filter(SITUACION_EGRESO == 1)

Adm2019 <- Adm2019 %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2019_grouped <- Adm2019 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    ) 

Adm2019_grouped <- Adm2019_grouped %>% 
  left_join(Adm2019 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")

```

## 2018 (where 0 are treated as NA)

```{r}
Adm2018_na <- Adm2018 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  )
Adm2018_na <- Adm2018_na %>%
  filter(SITUACION_EGRESO == 1)
Adm2018_na <- Adm2018_na %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)
Adm2018_na_grouped <- Adm2018_na %>%
  group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
    )

Adm2018_na_grouped <- Adm2018_na_grouped %>%
  left_join(Adm2018_na %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")
```


## 2018 (where 0 are counted in the calculations)

```{r}
Adm2018 <- Adm2018 %>% 
  filter(SITUACION_EGRESO == 1)

Adm2018 <- Adm2018 %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2018_grouped <- Adm2018 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    ) 

Adm2018_grouped <- Adm2018_grouped %>% 
  left_join(Adm2018 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")

```

## 2017 (where 0 are treated as NA)

```{r}
Adm2017_na <- Adm2017 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  )
Adm2017_na <- Adm2017_na %>%
  filter(SITUACION_EGRESO == 1)
Adm2017_na <- Adm2017_na %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION,
 CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)
Adm2017_na_grouped <- Adm2017_na %>%
  group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
    )
Adm2017_na_grouped <- Adm2017_na_grouped %>%
  left_join(Adm2017_na %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")

```



## 2017 (where 0 is counted in the calculations)

```{r}
Adm2017 <- Adm2017 %>% 
  filter(SITUACION_EGRESO == 1)

Adm2017 <- Adm2017 %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2017_grouped <- Adm2017 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    ) 

Adm2017_grouped <- Adm2017_grouped %>% 
  left_join(Adm2017 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")

```

## 2016 (where 0 are treated as NA)

```{r}
Adm2016_na <- Adm2016 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  )

Adm2016_na <- Adm2016_na %>%
  filter(SITUACION_EGRESO == 1)
Adm2016_na <- Adm2016_na %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)
Adm2016_na_grouped <- Adm2016_na %>%
  group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
    )
Adm2016_na_grouped <- Adm2016_na_grouped %>%
  left_join(Adm2016_na %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")


```

## 2016 (where 0 are counted in the calculations)

```{r}
Adm2016 <- Adm2016 %>% 
  filter(SITUACION_EGRESO == 1)

## I also just want the features, ID, RBD, GRUPO_DEPENDENCIA,CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM
Adm2016 <- Adm2016 %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

# i want make groups by RBD and compute the average of Leng_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM for each group.
Adm2016_grouped <- Adm2016 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    ) 
# I want add the column of GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA to the grouped data frame. I will use the RBD to join the original data frame with the grouped data frame. I will use the left_join function from the dplyr package to do this.
Adm2016_grouped <- Adm2016_grouped %>% 
  left_join(Adm2016 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")

```

## 2015 (where 0 is trated as NA)

```{r}
Adm2015_na <- Adm2015 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  )
Adm2015_na <- Adm2015_na %>%
  filter(SITUACION_EGRESO == 1)
Adm2015_na <- Adm2015_na %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)
Adm2015_na_grouped <- Adm2015_na %>%
    group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
    )
Adm2015_na_grouped <- Adm2015_na_grouped %>%
  left_join(Adm2015_na %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")
```



## 2015 (Where 0 is counted in the calculations)

```{r}
Adm2015 <- ArchivoC_Adm2015 %>% 
  filter(SITUACION_EGRESO == 1) %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2015_grouped <- Adm2015 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    )  %>% left_join(Adm2015 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")
```

## 2014 (where 0 is counted as NA)

```{r}
Adm2014_na <- Adm2014 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  )
Adm2014_na <- Adm2014_na %>%
  filter(SITUACION_EGRESO == 1)
Adm2014_na <- Adm2014_na %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)
Adm2014_na_grouped <- Adm2014_na %>%
    group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
    )
Adm2014_na_grouped <- Adm2014_na_grouped %>%
  left_join(Adm2014_na %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")
```


## 2014 (where 0 is counted in the calculations)

```{r}
Adm2014 <- ArchivoC_Adm2014 %>% 
  filter(SITUACION_EGRESO == 1) %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2014_grouped <- Adm2014 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    )  %>% left_join(Adm2014 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")

```

## 2013 (where 0 is treated as NA)

```{r}
Adm2013_na <- Adm2013 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  ) %>%
  filter(SITUACION_EGRESO == 1) %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA,
         LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2013_na_grouped <- Adm2013_na %>%   # ✅ fixed here
  group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
  )

Adm2013_na_grouped <- Adm2013_na_grouped %>%
  left_join(
    Adm2013_na %>%
      select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>%
      distinct(),
    by = "RBD"
  )
```


## 2013 (where 0 is counted in the calculations)

```{r}
Adm2013 <- ArchivoC_Adm2013 %>% 
  filter(SITUACION_EGRESO == 1) %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2013_grouped <- Adm2013 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    ) %>% left_join(Adm2013 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")
```

## 2012 (where 0 is treated as NA)'

```{r}
Adm2012_na <- Adm2012 %>%
  mutate(
    LENG_ACTUAL = na_if(LENG_ACTUAL, 0),
    MATE_ACTUAL = na_if(MATE_ACTUAL, 0),
    HCSO_ACTUAL = na_if(HCSO_ACTUAL, 0),
    CIEN_ACTUAL = na_if(CIEN_ACTUAL, 0),
    PTJE_NEM = na_if(PTJE_NEM, 0)
  ) %>%
  filter(SITUACION_EGRESO == 1) %>%
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA,
         LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2012_na_grouped <- Adm2012_na %>%   # ✅ fixed here
  group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_LENG = sum(!is.na(LENG_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
  )

Adm2012_na_grouped <- Adm2012_na_grouped %>%
  left_join(
    Adm2012_na %>%
      select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>%
      distinct(),
    by = "RBD"
  )
```

## 2012 (where 0 is counted in the calculations)
```{r}
Adm2012 <- ArchivoC_Adm2012 %>% 
  filter(SITUACION_EGRESO == 1) %>% 
  select(ID_aux, RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA, LENG_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, PTJE_NEM)

Adm2012_grouped <- Adm2012 %>% 
    group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(LENG_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    n_Students_RBD = n(),
    LENG_ACTUAL_p10 = quantile(LENG_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(LENG_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(LENG_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(LENG_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(LENG_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE)
    ) %>% left_join(Adm2012 %>% select(RBD, GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA) %>% distinct(), by = "RBD")
```

Now we have the grouped data frames for each year. We can combine them into a single data frame for analysis. We will use the bind_rows function from the dplyr package to do this. I give an additional column to each adm year to identify the year of the data.



## Adm_combined_na (0 treated as NA in the calculations)

```{r}
Adm2020_grouped_na <- Adm2020_na_grouped %>% mutate(year = 2020)
Adm2019_grouped_na <- Adm2019_na_grouped %>% mutate(year = 2019)
Adm2018_grouped_na <- Adm2018_na_grouped %>% mutate(year = 2018)
Adm2017_grouped_na <- Adm2017_na_grouped %>% mutate(year = 2017)
Adm2016_grouped_na <- Adm2016_na_grouped %>% mutate(year = 2016)
Adm2015_grouped_na <- Adm2015_na_grouped %>% mutate(year = 2015)
Adm2014_grouped_na <- Adm2014_na_grouped %>% mutate(year = 2014)
Adm2013_grouped_na <- Adm2013_na_grouped %>% mutate(year = 2013)
Adm2012_grouped_na <- Adm2012_na_grouped %>% mutate(year = 2012)
Adm_combined_na_2012_2020 <- bind_rows(Adm2020_grouped_na, Adm2019_grouped_na, Adm2018_grouped_na, Adm2017_grouped_na, Adm2016_grouped_na, Adm2015_grouped_na, Adm2014_grouped_na, Adm2013_grouped_na, Adm2012_grouped_na)
```




## Adm_combined (0 counted in the calculations)

```{r}
Adm2020_grouped <- Adm2020_grouped %>% mutate(year = 2020)
Adm2019_grouped <- Adm2019_grouped %>% mutate(year = 2019)
Adm2018_grouped <- Adm2018_grouped %>% mutate(year = 2018)
Adm2017_grouped <- Adm2017_grouped %>% mutate(year = 2017)
Adm2016_grouped <- Adm2016_grouped %>% mutate(year = 2016)
Adm2015_grouped <- Adm2015_grouped %>% mutate(year = 2015)
Adm2014_grouped <- Adm2014_grouped %>% mutate(year = 2014)
Adm2013_grouped <- Adm2013_grouped %>% mutate(year = 2013)
Adm2012_grouped <- Adm2012_grouped %>% mutate(year = 2012)
Adm_combined_2012_2020 <- bind_rows(Adm2020_grouped, Adm2019_grouped, Adm2018_grouped, Adm2017_grouped, Adm2016_grouped, Adm2015_grouped, Adm2014_grouped, Adm2013_grouped, Adm2012_grouped)
                                  
                                    
```

I want to make groups in Adm_combined where are grouped by rbd, and whe can see the year in ascendent order. We can use the group_by and arrange functions from the dplyr package to do this.


## ADM_COMBINED_GROUPED_NA (where 0 is treated as NA in the calculations)

```{r}
Adm_combined_grouped_na_2012_2020 <- Adm_combined_na_2012_2020 %>% 
  group_by(RBD) %>% 
  arrange(year)

# could you change the column year to the second column in the data frame? We can use the select function from the dplyr package to do this.
Adm_combined_grouped_na_2012_2020 <- Adm_combined_grouped_na_2012_2020 %>% 
  select(RBD, year, everything())

# i want drop the rbd = 0 from the data frame, we can use the filter function from the dplyr package to do this.
Adm_combined_grouped_na_2012_2020 <- Adm_combined_grouped_na_2012_2020 %>% 
  filter(RBD != 0)
#I want that the order is by the rbd increasing
Adm_combined_grouped_na_2012_2020 <- Adm_combined_grouped_na_2012_2020 %>% 
  arrange(RBD)


#move the columns of n_student_ (LENG, MATE, HCSO, CIEN) to the end of the data frame with also GRUPO_DEPENDENCIA, CODIGO_REGION, CODIGO_COMUNA
Adm_combined_grouped_na_2012_2020 <- Adm_combined_grouped_na_2012_2020 %>% 
  select(-starts_with("n_Students_"), -GRUPO_DEPENDENCIA, -CODIGO_REGION, -CODIGO_COMUNA, everything())


View(Adm_combined_grouped_na_2012_2020)
```


## ADM_COMBINED_GROUPED (where 0 is counted in the calculations)

```{r}
Adm_combined_grouped_2012_2020 <- Adm_combined_2012_2020 %>% 
  group_by(RBD) %>% 
  arrange(year)

# could you change the column year to the second column in the data frame? We can use the select function from the dplyr package to do this.

Adm_combined_grouped_2012_2020 <- Adm_combined_grouped_2012_2020 %>% 
  select(RBD, year, everything())

# i want drop the rbd = 0 from the data frame, we can use the filter function from the dplyr package to do this.

Adm_combined_grouped_2012_2020 <- Adm_combined_grouped_2012_2020 %>% 
  filter(RBD != 0)

#I want that the order is by the rbd increasing
Adm_combined_grouped_2012_2020 <- Adm_combined_grouped_2012_2020 %>% 
  arrange(RBD)


# move 'n_Students' change the name of the column to N_Students_RBD. And then move the column "N_Studnets_RBD" to the end of the dataframe
Adm_combined_grouped_2012_2020 <- Adm_combined_grouped_2012_2020 %>% 
  rename(N_Students_RBD = n_Students) %>%
  select(-starts_with("n_Students_"), everything())


view(Adm_combined_grouped_2012_2020)


```

```{r}
#Exporting csv where 0 are counted in the calculations
write.csv(Adm_combined_grouped_2012_2020, "Adm_combined_grouped_2012_2020.csv", row.names = FALSE)

#Exporting csv where 0 are treated as NA in the calculations
write.csv(Adm_combined_grouped_na_2012_2020, "Adm_combined_grouped_na_2012_2020.csv", row.names = FALSE)

```
NOTA: por alguna razon en rbd 2013 (n_de estudiantes) no son iguales
NOTA2: Arreglado :)

## Reading 2021 - 2025

```{r}
# I want to create a new data frame for the years 2021 to 2025, we can use the filter function from the dplyr package to do this.
ArchivoC_Adm2021 <- read_csv("Data/A_INSCRITOS_PUNTAJES_PDT_2021_PUB_MRUN.csv")
ArchivoC_Adm2022 <- read_csv("Data/A_INSCRITOS_PUNTAJES_PDT_2022_PUB_MRUN.csv")
ArchivoC_Adm2023 <- read_csv("Data/A_INSCRITOS_PUNTAJES_2023_PAES_PUB_MRUN.csv")
ArchivoC_Adm2024 <- read_csv("Data/A_INSCRITOS_PUNTAJES_PAES_2024_PUB_MRUN.csv")
ArchivoC_Adm2025 <- read_csv("Data/A_INSCRITOS_PUNTAJES_PAES_2025_PUB_MRUN.csv")

```

## Delimiter 2:

```{r}
Adm2021 <- ArchivoC_Adm2021 <- read_delim("Data/A_INSCRITOS_PUNTAJES_PDT_2021_PUB_MRUN.csv", 
                               delim = ";")
Adm2022 <- ArchivoC_Adm2022 <- read_delim("Data/A_INSCRITOS_PUNTAJES_PDT_2022_PUB_MRUN.csv", 
                               delim = ";")
Adm2023 <- ArchivoC_Adm2023 <- read_delim("Data/A_INSCRITOS_PUNTAJES_2023_PAES_PUB_MRUN.csv", 
                               delim = ";")
Adm2024 <- ArchivoC_Adm2024 <- read_delim("Data/A_INSCRITOS_PUNTAJES_PAES_2024_PUB_MRUN.csv", 
                               delim = ";")
Adm2025 <- ArchivoC_Adm2025 <- read_delim("Data/A_INSCRITOS_PUNTAJES_PAES_2025_PUB_MRUN.csv", 
                               delim = ";")
```

## 2021 (where 0 is counted in the calculations)

```{r}
## 2021
Adm2021 <- Adm2021 %>% 
  filter(ANYO_DE_EGRESO == 2020, 
         PUNTAJES_PROCESO == 1, 
         TIPO_RENDICION == 1) %>% 
  select(ANYO_PROCESO, RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, 
         CODIGO_COMUNA_EGRESO, PUNTAJES_PROCESO, 
         CLEC_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, 
         PTJE_NEM, ANYO_DE_EGRESO, TIPO_RENDICION) %>% 
  mutate(
    CLEC_ACTUAL = as.numeric(gsub(",", ".", CLEC_ACTUAL)),
    MATE_ACTUAL = as.numeric(gsub(",", ".", MATE_ACTUAL)),
    HCSO_ACTUAL = as.numeric(gsub(",", ".", HCSO_ACTUAL)),
    CIEN_ACTUAL = as.numeric(gsub(",", ".", CIEN_ACTUAL)),
    PTJE_NEM = as.numeric(PTJE_NEM)
  )

Adm2021_grouped <- Adm2021 %>% 
  group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(CLEC_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(CLEC_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(CLEC_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(CLEC_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(CLEC_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(CLEC_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_RBD= n()
    ) %>%
  left_join(
    Adm2021 %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2021_grouped)
```
## 2021 (where 0 is treated as NA in the calculations)



```{r}
## 2021 (same filters as Adm2021, but zeros -> NA)
Adm2021_na <- Adm2021 %>%   # use your raw dataset here
  filter(
    ANYO_DE_EGRESO == 2020,
    PUNTAJES_PROCESO == 1,
    TIPO_RENDICION == 1
  ) %>%
  select(
    ANYO_PROCESO, RBD, DEPENDENCIA, CODIGO_REGION_EGRESO,
    CODIGO_COMUNA_EGRESO, PUNTAJES_PROCESO,
    CLEC_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL,
    PTJE_NEM, ANYO_DE_EGRESO, TIPO_RENDICION
  ) %>%
  mutate(
    CLEC_ACTUAL = na_if(as.numeric(gsub(",", ".", CLEC_ACTUAL)), 0),
    MATE_ACTUAL = na_if(as.numeric(gsub(",", ".", MATE_ACTUAL)), 0),
    HCSO_ACTUAL = na_if(as.numeric(gsub(",", ".", HCSO_ACTUAL)), 0),
    CIEN_ACTUAL = na_if(as.numeric(gsub(",", ".", CIEN_ACTUAL)), 0),
    PTJE_NEM = na_if(as.numeric(PTJE_NEM), 0)
  )

Adm2021_na_grouped <- Adm2021_na %>%
  group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(CLEC_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    LENG_ACTUAL_p10 = quantile(CLEC_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(CLEC_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(CLEC_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(CLEC_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(CLEC_ACTUAL, 0.90, na.rm = TRUE),

    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),

    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),

    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),

    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),

    n_Students_LENG = sum(!is.na(CLEC_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
  ) %>%
  left_join(
    Adm2021_na %>%
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>%
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2021_na_grouped)

```


## 2022 (where 0 is counted in the calculations

```{r}
## 2022
Adm2022 <- Adm2022 %>% 
  filter(ANYO_DE_EGRESO == 2021, 
         PUNTAJES_PROCESO == 1 
         ) %>% 
  select(ANYO_PROCESO, RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, 
         CODIGO_COMUNA_EGRESO, PUNTAJES_PROCESO, 
         CLEC_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL, 
         PTJE_NEM, ANYO_DE_EGRESO) %>% 
  mutate(
    CLEC_ACTUAL = as.numeric(gsub(",", ".", CLEC_ACTUAL)),
    MATE_ACTUAL = as.numeric(gsub(",", ".", MATE_ACTUAL)),
    HCSO_ACTUAL = as.numeric(gsub(",", ".", HCSO_ACTUAL)),
    CIEN_ACTUAL = as.numeric(gsub(",", ".", CIEN_ACTUAL)),
    PTJE_NEM = as.numeric(PTJE_NEM)
  )

Adm2022_grouped <- Adm2022 %>% 
  group_by(RBD) %>% 
  summarise(
    avg_LENG_ACTUAL = mean(CLEC_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),
    LENG_ACTUAL_p10 = quantile(CLEC_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(CLEC_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(CLEC_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(CLEC_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(CLEC_ACTUAL, 0.90, na.rm = TRUE),
    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),
    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),
    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    n_Students_RBD= n()
    ) %>%
  left_join(
    Adm2021 %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2022_grouped)
```


## 2022 (0 treated as NA)

```{r}
Adm2022_na <- Adm2022 %>%   # use your raw dataset here
  filter(
    ANYO_DE_EGRESO == 2021,
    PUNTAJES_PROCESO == 1
  ) %>%
  select(
    ANYO_PROCESO, RBD, DEPENDENCIA, CODIGO_REGION_EGRESO,
    CODIGO_COMUNA_EGRESO, PUNTAJES_PROCESO,
    CLEC_ACTUAL, MATE_ACTUAL, HCSO_ACTUAL, CIEN_ACTUAL,
    PTJE_NEM, ANYO_DE_EGRESO
  ) %>%
  mutate(
    CLEC_ACTUAL = na_if(as.numeric(gsub(",", ".", CLEC_ACTUAL)), 0),
    MATE_ACTUAL = na_if(as.numeric(gsub(",", ".", MATE_ACTUAL)), 0),
    HCSO_ACTUAL = na_if(as.numeric(gsub(",", ".", HCSO_ACTUAL)), 0),
    CIEN_ACTUAL = na_if(as.numeric(gsub(",", ".", CIEN_ACTUAL)), 0),
    PTJE_NEM = na_if(as.numeric(PTJE_NEM), 0)
  )

Adm2022_na_grouped <- Adm2022_na %>%
  group_by(RBD) %>%
  summarise(
    avg_LENG_ACTUAL = mean(CLEC_ACTUAL, na.rm = TRUE),
    avg_MATE_ACTUAL = mean(MATE_ACTUAL, na.rm = TRUE),
    avg_HCSO_ACTUAL = mean(HCSO_ACTUAL, na.rm = TRUE),
    avg_CIEN_ACTUAL = mean(CIEN_ACTUAL, na.rm = TRUE),
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    LENG_ACTUAL_p10 = quantile(CLEC_ACTUAL, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25 = quantile(CLEC_ACTUAL, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN = quantile(CLEC_ACTUAL, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75 = quantile(CLEC_ACTUAL, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90 = quantile(CLEC_ACTUAL, 0.90, na.rm = TRUE),

    MATE_ACTUAL_p10 = quantile(MATE_ACTUAL, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25 = quantile(MATE_ACTUAL, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN = quantile(MATE_ACTUAL, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75 = quantile(MATE_ACTUAL, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90 = quantile(MATE_ACTUAL, 0.90, na.rm = TRUE),

    HCSO_ACTUAL_p10 = quantile(HCSO_ACTUAL, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25 = quantile(HCSO_ACTUAL, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN = quantile(HCSO_ACTUAL, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75 = quantile(HCSO_ACTUAL, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90 = quantile(HCSO_ACTUAL, 0.90, na.rm = TRUE),

    CIEN_ACTUAL_p10 = quantile(CIEN_ACTUAL, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25 = quantile(CIEN_ACTUAL, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN = quantile(CIEN_ACTUAL, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75 = quantile(CIEN_ACTUAL, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90 = quantile(CIEN_ACTUAL, 0.90, na.rm = TRUE),

    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),

    n_Students_LENG = sum(!is.na(CLEC_ACTUAL)),
    n_Students_MATE = sum(!is.na(MATE_ACTUAL)),
    n_Students_HCSO = sum(!is.na(HCSO_ACTUAL)),
    n_Students_CIEN = sum(!is.na(CIEN_ACTUAL)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD = n()
  ) %>%
  left_join(
    Adm2022_na %>%
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>%
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2022_na_grouped)
```




## 2023 (0 counted in the calculations)

```{r}
# 1️⃣ Filter and select relevant columns
Adm2023 <- Adm2023 %>% 
  filter(
    ANYO_PROCESO == 2023, 
    RINDIO_PROCESO_ANTERIOR == 0) %>%
  select(
    ANYO_PROCESO, RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO,
    CLEC_MAX, MATE1_MAX, MATE2_MAX, HCSOC_MAX, CIEN_MAX, PTJE_NEM
  )

Adm2023_grouped <- Adm2023 %>% 
  group_by(RBD) %>% 
  summarise(
    ## --- Averages ---  ### From 2023 all the ACTUAL ARE ACTUALLY MAX SCORES
    avg_LENG_ACTUAL   = mean(CLEC_MAX, na.rm = TRUE),       ## MAX CLEC
    avg_MATE_ACTUAL  = mean(MATE1_MAX, na.rm = TRUE),       ## MAX MATH1
    avg_MATE2_MAX  = mean(MATE2_MAX, na.rm = TRUE),         ## Added for Matemática 2
    avg_HCSO_ACTUAL  = mean(HCSOC_MAX, na.rm = TRUE),       ## MAX HCSOC
    avg_CIEN_ACTUAL   = mean(CIEN_MAX, na.rm = TRUE),      ## MAX CIEN
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    ## --- Percentiles: Lenguaje ---
    LENG_ACTUAL_p10   = quantile(CLEC_MAX, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25   = quantile(CLEC_MAX, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN= quantile(CLEC_MAX, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75   = quantile(CLEC_MAX, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90   = quantile(CLEC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 1 ---
    MATE_ACTUAL_p10   = quantile(MATE1_MAX, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25   = quantile(MATE1_MAX, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN= quantile(MATE1_MAX, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75   = quantile(MATE1_MAX, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90   = quantile(MATE1_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 2 ---
    MATE2_MAX_p10   = quantile(MATE2_MAX, 0.10, na.rm = TRUE),
    MATE2_MAX_p25   = quantile(MATE2_MAX, 0.25, na.rm = TRUE),
    MATE2_MAX_MEDIAN= quantile(MATE2_MAX, 0.50, na.rm = TRUE),
    MATE2_MAX_p75   = quantile(MATE2_MAX, 0.75, na.rm = TRUE),
    MATE2_MAX_p90   = quantile(MATE2_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Historia y Ciencias Sociales ---
    HCSO_ACTUAL_p10   = quantile(HCSOC_MAX, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25   = quantile(HCSOC_MAX, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN= quantile(HCSOC_MAX, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75   = quantile(HCSOC_MAX, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90   = quantile(HCSOC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Ciencias ---
    CIEN_ACTUAL_p10   = quantile(CIEN_MAX, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25   = quantile(CIEN_MAX, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN= quantile(CIEN_MAX, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75   = quantile(CIEN_MAX, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90   = quantile(CIEN_MAX, 0.90, na.rm = TRUE),
    ## --- Percentiles: Puntaje Nem
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),

    ## --- Student counts per subject ---
    n_Students_RBD   = n()
  ) %>%
  ungroup() %>%
  left_join(
    Adm2023 %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2023_grouped)


```




## --- 2023 (0 treated as NA) ---
```{r}

# 2️⃣ Convert scores to numeric and replace 0 with NA
Adm2023_na <- Adm2023 %>% 
  mutate(
    CLEC_MAX   = na_if(as.numeric(gsub(",", ".", CLEC_MAX)), 0),
    MATE1_MAX  = na_if(as.numeric(gsub(",", ".", MATE1_MAX)), 0),
    MATE2_MAX  = na_if(as.numeric(gsub(",", ".", MATE2_MAX)), 0),
    HCSOC_MAX  = na_if(as.numeric(gsub(",", ".", HCSOC_MAX)), 0),
    CIEN_MAX   = na_if(as.numeric(gsub(",", ".", CIEN_MAX)), 0),
    PTJE_NEM = na_if(as.numeric(PTJE_NEM), 0)
  )

# 3️⃣ Group by RBD and calculate summary statistics
Adm2023_na_grouped <- Adm2023_na %>% 
  group_by(RBD) %>% 
  summarise(
    ## --- Averages ---
    ## --- Averages ---  ### From 2023 all the ACTUAL ARE ACTUALLY MAX SCORES
    avg_LENG_ACTUAL   = mean(CLEC_MAX, na.rm = TRUE),       ## MAX CLEC
    avg_MATE_ACTUAL  = mean(MATE1_MAX, na.rm = TRUE),       ## MAX MATH1
    avg_MATE2_MAX  = mean(MATE2_MAX, na.rm = TRUE),         ## Added for Matemática 2
    avg_HCSO_ACTUAL  = mean(HCSOC_MAX, na.rm = TRUE),       ## MAX HCSOC
    avg_CIEN_ACTUAL   = mean(CIEN_MAX, na.rm = TRUE),    ## MAX CIEN
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    ## --- Percentiles: Lenguaje ---
    LENG_ACTUAL_p10   = quantile(CLEC_MAX, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25   = quantile(CLEC_MAX, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN= quantile(CLEC_MAX, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75   = quantile(CLEC_MAX, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90   = quantile(CLEC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 1 ---
    MATE_ACTUAL_p10   = quantile(MATE1_MAX, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25   = quantile(MATE1_MAX, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN= quantile(MATE1_MAX, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75   = quantile(MATE1_MAX, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90   = quantile(MATE1_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 2 ---
    MATE2_MAX_p10   = quantile(MATE2_MAX, 0.10, na.rm = TRUE),
    MATE2_MAX_p25   = quantile(MATE2_MAX, 0.25, na.rm = TRUE),
    MATE2_MAX_MEDIAN= quantile(MATE2_MAX, 0.50, na.rm = TRUE),
    MATE2_MAX_p75   = quantile(MATE2_MAX, 0.75, na.rm = TRUE),
    MATE2_MAX_p90   = quantile(MATE2_MAX, 0.90, na.rm = TRUE),

     ## --- Percentiles: Historia y Ciencias Sociales ---
    HCSO_ACTUAL_p10   = quantile(HCSOC_MAX, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25   = quantile(HCSOC_MAX, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN= quantile(HCSOC_MAX, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75   = quantile(HCSOC_MAX, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90   = quantile(HCSOC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Ciencias ---
    CIEN_ACTUAL_p10   = quantile(CIEN_MAX, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25   = quantile(CIEN_MAX, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN= quantile(CIEN_MAX, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75   = quantile(CIEN_MAX, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90   = quantile(CIEN_MAX, 0.90, na.rm = TRUE),
    
    ## --- Percentiles: Puntaje Nem
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    
    ## --- Student counts per subject ---
    n_Students_LENG  = sum(!is.na(CLEC_MAX)),
    n_Students_MATE = sum(!is.na(MATE1_MAX)),
    n_Students_MATE2 = sum(!is.na(MATE2_MAX)),
    n_Students_HCSO  = sum(!is.na(HCSOC_MAX)),
    n_Students_CIEN  = sum(!is.na(CIEN_MAX)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD   = n()
  ) %>%
  ungroup() %>%
  left_join(
    Adm2023_na %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2023_na_grouped)
```


## 2024 (0 counted in the calculations)

```{r}
# 1️⃣ Filter and select relevant columns
Adm2024 <- Adm2024 %>% 
  filter(
    ANYO_PROCESO == 2024, 
    RINDIO_PROCESO_ANTERIOR == 0) %>%
  select(
    ANYO_PROCESO, RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO,
    CLEC_MAX, MATE1_MAX, MATE2_MAX, HCSOC_MAX, CIEN_MAX, PTJE_NEM
  )

Adm2024_grouped <- Adm2024 %>% 
  group_by(RBD) %>% 
  summarise(
    ## --- Averages ---
    ## --- Averages ---  ### From 2023 all the ACTUAL ARE ACTUALLY MAX SCORES
    avg_LENG_ACTUAL   = mean(CLEC_MAX, na.rm = TRUE),       ## MAX CLEC
    avg_MATE_ACTUAL  = mean(MATE1_MAX, na.rm = TRUE),       ## MAX MATH1
    avg_MATE2_MAX  = mean(MATE2_MAX, na.rm = TRUE),         ## Added for Matemática 2
    avg_HCSO_ACTUAL  = mean(HCSOC_MAX, na.rm = TRUE),       ## MAX HCSOC
    avg_CIEN_ACTUAL   = mean(CIEN_MAX, na.rm = TRUE),        ## MAX CIEN
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    ## --- Percentiles: Lenguaje ---
    LENG_ACTUAL_p10   = quantile(CLEC_MAX, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25   = quantile(CLEC_MAX, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN= quantile(CLEC_MAX, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75   = quantile(CLEC_MAX, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90   = quantile(CLEC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 1 ---
    MATE_ACTUAL_p10   = quantile(MATE1_MAX, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25   = quantile(MATE1_MAX, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN= quantile(MATE1_MAX, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75   = quantile(MATE1_MAX, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90   = quantile(MATE1_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 2 ---
    MATE2_MAX_p10   = quantile(MATE2_MAX, 0.10, na.rm = TRUE),
    MATE2_MAX_p25   = quantile(MATE2_MAX, 0.25, na.rm = TRUE),
    MATE2_MAX_MEDIAN= quantile(MATE2_MAX, 0.50, na.rm = TRUE),
    MATE2_MAX_p75   = quantile(MATE2_MAX, 0.75, na.rm = TRUE),
    MATE2_MAX_p90   = quantile(MATE2_MAX, 0.90, na.rm = TRUE),

     ## --- Percentiles: Historia y Ciencias Sociales ---
    HCSO_ACTUAL_p10   = quantile(HCSOC_MAX, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25   = quantile(HCSOC_MAX, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN= quantile(HCSOC_MAX, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75   = quantile(HCSOC_MAX, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90   = quantile(HCSOC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Ciencias ---
    CIEN_ACTUAL_p10   = quantile(CIEN_MAX, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25   = quantile(CIEN_MAX, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN= quantile(CIEN_MAX, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75   = quantile(CIEN_MAX, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90   = quantile(CIEN_MAX, 0.90, na.rm = TRUE),
    ## --- Percentiles: Puntaje Nem
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    
    ## --- Student counts per subject ---
    n_Students_RBD   = n()
  ) %>%
  ungroup() %>%
  left_join(
    Adm2024 %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2024_grouped)
```

## --- 2024 (0 treated as NA) ---
```{r}

# 2️⃣ Convert scores to numeric and replace 0 with NA
Adm2024_na <- Adm2024 %>% 
  mutate(
    CLEC_MAX   = na_if(as.numeric(gsub(",", ".", CLEC_MAX)), 0),
    MATE1_MAX  = na_if(as.numeric(gsub(",", ".", MATE1_MAX)), 0),
    MATE2_MAX  = na_if(as.numeric(gsub(",", ".", MATE2_MAX)), 0),
    HCSOC_MAX  = na_if(as.numeric(gsub(",", ".", HCSOC_MAX)), 0),
    CIEN_MAX   = na_if(as.numeric(gsub(",", ".", CIEN_MAX)), 0),
    PTJE_NEM = na_if(as.numeric(PTJE_NEM), 0)
  )

# 3️⃣ Group by RBD and calculate summary statistics
Adm2024_na_grouped <- Adm2024_na %>% 
  group_by(RBD) %>% 
  summarise(
    ## --- Averages ---
    ## --- Averages ---  ### From 2023 all the ACTUAL ARE ACTUALLY MAX SCORES
    avg_LENG_ACTUAL   = mean(CLEC_MAX, na.rm = TRUE),       ## MAX CLEC
    avg_MATE_ACTUAL  = mean(MATE1_MAX, na.rm = TRUE),       ## MAX MATH1
    avg_MATE2_MAX  = mean(MATE2_MAX, na.rm = TRUE),         ## Added for Matemática 2
    avg_HCSO_ACTUAL  = mean(HCSOC_MAX, na.rm = TRUE),       ## MAX HCSOC
    avg_CIEN_ACTUAL   = mean(CIEN_MAX, na.rm = TRUE),        ## MAX CIEN
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    ## --- Percentiles: Lenguaje ---
    LENG_ACTUAL_p10   = quantile(CLEC_MAX, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25   = quantile(CLEC_MAX, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN= quantile(CLEC_MAX, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75   = quantile(CLEC_MAX, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90   = quantile(CLEC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 1 ---
    MATE_ACTUAL_p10   = quantile(MATE1_MAX, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25   = quantile(MATE1_MAX, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN= quantile(MATE1_MAX, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75   = quantile(MATE1_MAX, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90   = quantile(MATE1_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 2 ---
    MATE2_MAX_p10   = quantile(MATE2_MAX, 0.10, na.rm = TRUE),
    MATE2_MAX_p25   = quantile(MATE2_MAX, 0.25, na.rm = TRUE),
    MATE2_MAX_MEDIAN= quantile(MATE2_MAX, 0.50, na.rm = TRUE),
    MATE2_MAX_p75   = quantile(MATE2_MAX, 0.75, na.rm = TRUE),
    MATE2_MAX_p90   = quantile(MATE2_MAX, 0.90, na.rm = TRUE),
    
     ## --- Percentiles: Historia y Ciencias Sociales ---
    HCSO_ACTUAL_p10   = quantile(HCSOC_MAX, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25   = quantile(HCSOC_MAX, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN= quantile(HCSOC_MAX, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75   = quantile(HCSOC_MAX, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90   = quantile(HCSOC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Ciencias ---
    CIEN_ACTUAL_p10   = quantile(CIEN_MAX, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25   = quantile(CIEN_MAX, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN= quantile(CIEN_MAX, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75   = quantile(CIEN_MAX, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90   = quantile(CIEN_MAX, 0.90, na.rm = TRUE),
    ## --- Percentiles: Puntaje Nem
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    ## --- Student counts per subject ---
    n_Students_LENG  = sum(!is.na(CLEC_MAX)),
    n_Students_MATE = sum(!is.na(MATE1_MAX)),
    n_Students_MATE2 = sum(!is.na(MATE2_MAX)),
    n_Students_HCSO  = sum(!is.na(HCSOC_MAX)),
    n_Students_CIEN  = sum(!is.na(CIEN_MAX)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD   = n()
  ) %>%
  ungroup() %>%
  left_join(
    Adm2024_na %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2024_na_grouped)
```



## 2025 (0 counted in the calculations)

```{r}
# 1️⃣ Filter and select relevant columns
Adm2025 <- Adm2025 %>% 
  filter(
    ANYO_PROCESO == 2025, 
    RINDIO_PROCESO_ANTERIOR == 0) %>%
  select(
    ANYO_PROCESO, RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO,
    CLEC_MAX, MATE1_MAX, MATE2_MAX, HCSOC_MAX, CIEN_MAX, PTJE_NEM
  )

Adm2025_grouped <- Adm2025 %>% 
  group_by(RBD) %>% 
  summarise(
    ## --- Averages ---
    ## --- Averages ---  ### From 2023 all the ACTUAL ARE ACTUALLY MAX SCORES
    avg_LENG_ACTUAL   = mean(CLEC_MAX, na.rm = TRUE),       ## MAX CLEC
    avg_MATE_ACTUAL  = mean(MATE1_MAX, na.rm = TRUE),       ## MAX MATH1
    avg_MATE2_MAX  = mean(MATE2_MAX, na.rm = TRUE),         ## Added for Matemática 2
    avg_HCSO_ACTUAL  = mean(HCSOC_MAX, na.rm = TRUE),       ## MAX HCSOC
    avg_CIEN_ACTUAL   = mean(CIEN_MAX, na.rm = TRUE),        ## MAX CIEN
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    ## --- Percentiles: Lenguaje ---
    LENG_ACTUAL_p10   = quantile(CLEC_MAX, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25   = quantile(CLEC_MAX, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN= quantile(CLEC_MAX, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75   = quantile(CLEC_MAX, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90   = quantile(CLEC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 1 ---
    MATE_ACTUAL_p10   = quantile(MATE1_MAX, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25   = quantile(MATE1_MAX, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN= quantile(MATE1_MAX, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75   = quantile(MATE1_MAX, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90   = quantile(MATE1_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 2 ---
    MATE2_MAX_p10   = quantile(MATE2_MAX, 0.10, na.rm = TRUE),
    MATE2_MAX_p25   = quantile(MATE2_MAX, 0.25, na.rm = TRUE),
    MATE2_MAX_MEDIAN= quantile(MATE2_MAX, 0.50, na.rm = TRUE),
    MATE2_MAX_p75   = quantile(MATE2_MAX, 0.75, na.rm = TRUE),
    MATE2_MAX_p90   = quantile(MATE2_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Historia y Ciencias Sociales ---
    HCSO_ACTUAL_p10   = quantile(HCSOC_MAX, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25   = quantile(HCSOC_MAX, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN= quantile(HCSOC_MAX, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75   = quantile(HCSOC_MAX, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90   = quantile(HCSOC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Ciencias ---
    CIEN_ACTUAL_p10   = quantile(CIEN_MAX, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25   = quantile(CIEN_MAX, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN= quantile(CIEN_MAX, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75   = quantile(CIEN_MAX, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90   = quantile(CIEN_MAX, 0.90, na.rm = TRUE),
    ## --- Percentiles: Puntaje Nem
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    ## --- Student counts per subject ---
    n_Students_RBD   = n()
  ) %>%
  ungroup() %>%
  left_join(
    Adm2025 %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2025_grouped)
```

## --- 2025 (0 treated as NA) ---
```{r}

# 2️⃣ Convert scores to numeric and replace 0 with NA
Adm2025_na <- Adm2025 %>% 
  mutate(
    CLEC_MAX   = na_if(as.numeric(gsub(",", ".", CLEC_MAX)), 0),
    MATE1_MAX  = na_if(as.numeric(gsub(",", ".", MATE1_MAX)), 0),
    MATE2_MAX  = na_if(as.numeric(gsub(",", ".", MATE2_MAX)), 0),
    HCSO_MAX  = na_if(as.numeric(gsub(",", ".", HCSOC_MAX)), 0),
    CIEN_MAX   = na_if(as.numeric(gsub(",", ".", CIEN_MAX)), 0),
    PTJE_NEM = na_if(as.numeric(PTJE_NEM), 0)
  )

# 3️⃣ Group by RBD and calculate summary statistics
Adm2025_na_grouped <- Adm2025_na %>% 
  group_by(RBD) %>% 
  summarise(
    ## --- Averages ---
    ## --- Averages ---  ### From 2023 all the ACTUAL ARE ACTUALLY MAX SCORES
    avg_LENG_ACTUAL   = mean(CLEC_MAX, na.rm = TRUE),       ## MAX CLEC
    avg_MATE_ACTUAL  = mean(MATE1_MAX, na.rm = TRUE),       ## MAX MATH1
    avg_MATE2_MAX  = mean(MATE2_MAX, na.rm = TRUE),         ## Added for Matemática 2
    avg_HCSO_ACTUAL  = mean(HCSOC_MAX, na.rm = TRUE),       ## MAX HCSOC
    avg_CIEN_ACTUAL   = mean(CIEN_MAX, na.rm = TRUE),        ## MAX CIEN
    avg_PTJE_NEM = mean(PTJE_NEM, na.rm = TRUE),

    ## --- Percentiles: Lenguaje ---
    LENG_ACTUAL_p10   = quantile(CLEC_MAX, 0.10, na.rm = TRUE),
    LENG_ACTUAL_p25   = quantile(CLEC_MAX, 0.25, na.rm = TRUE),
    LENG_ACTUAL_MEDIAN= quantile(CLEC_MAX, 0.50, na.rm = TRUE),
    LENG_ACTUAL_p75   = quantile(CLEC_MAX, 0.75, na.rm = TRUE),
    LENG_ACTUAL_p90   = quantile(CLEC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 1 ---
    MATE_ACTUAL_p10   = quantile(MATE1_MAX, 0.10, na.rm = TRUE),
    MATE_ACTUAL_p25   = quantile(MATE1_MAX, 0.25, na.rm = TRUE),
    MATE_ACTUAL_MEDIAN= quantile(MATE1_MAX, 0.50, na.rm = TRUE),
    MATE_ACTUAL_p75   = quantile(MATE1_MAX, 0.75, na.rm = TRUE),
    MATE_ACTUAL_p90   = quantile(MATE1_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Matemática 2 ---
    MATE2_MAX_p10   = quantile(MATE2_MAX, 0.10, na.rm = TRUE),
    MATE2_MAX_p25   = quantile(MATE2_MAX, 0.25, na.rm = TRUE),
    MATE2_MAX_MEDIAN= quantile(MATE2_MAX, 0.50, na.rm = TRUE),
    MATE2_MAX_p75   = quantile(MATE2_MAX, 0.75, na.rm = TRUE),
    MATE2_MAX_p90   = quantile(MATE2_MAX, 0.90, na.rm = TRUE),

     ## --- Percentiles: Historia y Ciencias Sociales ---
    HCSO_ACTUAL_p10   = quantile(HCSOC_MAX, 0.10, na.rm = TRUE),
    HCSO_ACTUAL_p25   = quantile(HCSOC_MAX, 0.25, na.rm = TRUE),
    HCSO_ACTUAL_MEDIAN= quantile(HCSOC_MAX, 0.50, na.rm = TRUE),
    HCSO_ACTUAL_p75   = quantile(HCSOC_MAX, 0.75, na.rm = TRUE),
    HCSO_ACTUAL_p90   = quantile(HCSOC_MAX, 0.90, na.rm = TRUE),

    ## --- Percentiles: Ciencias ---
    CIEN_ACTUAL_p10   = quantile(CIEN_MAX, 0.10, na.rm = TRUE),
    CIEN_ACTUAL_p25   = quantile(CIEN_MAX, 0.25, na.rm = TRUE),
    CIEN_ACTUAL_MEDIAN= quantile(CIEN_MAX, 0.50, na.rm = TRUE),
    CIEN_ACTUAL_p75   = quantile(CIEN_MAX, 0.75, na.rm = TRUE),
    CIEN_ACTUAL_p90   = quantile(CIEN_MAX, 0.90, na.rm = TRUE),
    
    ## --- Percentiles: Puntaje Nem
    PTJE_NEM_p10 = quantile(PTJE_NEM, 0.10, na.rm = TRUE),
    PTJE_NEM_p25 = quantile(PTJE_NEM, 0.25, na.rm = TRUE),
    PTJE_NEM_MEDIAN = quantile(PTJE_NEM, 0.50, na.rm = TRUE),
    PTJE_NEM_p75 = quantile(PTJE_NEM, 0.75, na.rm = TRUE),
    PTJE_NEM_p90 = quantile(PTJE_NEM, 0.90, na.rm = TRUE),
    
    ## --- Student counts per subject ---
    n_Students_LENG  = sum(!is.na(CLEC_MAX)),
    n_Students_MATE = sum(!is.na(MATE1_MAX)),
    n_Students_MATE2 = sum(!is.na(MATE2_MAX)),
    n_Students_HCSO  = sum(!is.na(HCSOC_MAX)),
    n_Students_CIEN  = sum(!is.na(CIEN_MAX)),
    n_Students_PTJE_NEM = sum(!is.na(PTJE_NEM)),
    n_Students_RBD   = n()
  ) %>%
  ungroup() %>%
  left_join(
    Adm2025_na %>% 
      select(RBD, DEPENDENCIA, CODIGO_REGION_EGRESO, CODIGO_COMUNA_EGRESO) %>% 
      distinct(),
    by = "RBD"
  ) %>%
  rename(
    GRUPO_DEPENDENCIA = DEPENDENCIA,
    CODIGO_REGION = CODIGO_REGION_EGRESO,
    CODIGO_COMUNA = CODIGO_COMUNA_EGRESO
  )

View(Adm2025_na_grouped)
```


```{r}
# Ensure consistent types before combining
Adm2021_grouped <- Adm2021_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2022_grouped <- Adm2022_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2023_grouped <- Adm2023_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2024_grouped <- Adm2024_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2025_grouped <- Adm2025_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )


Adm2025_grouped <- Adm2025_grouped %>% mutate(year = 2025)
Adm2024_grouped <- Adm2024_grouped %>% mutate(year = 2024)
Adm2023_grouped <- Adm2023_grouped %>% mutate(year = 2023)
Adm2022_grouped <- Adm2022_grouped %>% mutate(year = 2022)
Adm2021_grouped <- Adm2021_grouped %>% mutate(year = 2021)

Adm_combined_2021_2025 <- bind_rows(
  Adm2025_grouped,
  Adm2024_grouped, 
  Adm2023_grouped, 
  Adm2022_grouped, 
  Adm2021_grouped
)
Adm_combined_21_25 <- Adm_combined_2021_2025 %>% 
  mutate(RBD = trimws(RBD)) %>%        # remove leading/trailing spaces
  filter(!is.na(RBD), RBD != "")


Adm_combined_21_25 <- Adm_combined_21_25 %>%
  select(
    year, everything()) %>%
  arrange(RBD, year)
View(Adm_combined_21_25)
```




```{r}
Adm2021_na_grouped <- Adm2021_na_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2022_na_grouped <- Adm2022_na_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2023_na_grouped <- Adm2023_na_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2024_na_grouped <- Adm2024_na_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )

Adm2025_na_grouped <- Adm2025_na_grouped %>%
  mutate(
    CODIGO_REGION = as.character(CODIGO_REGION),
    CODIGO_COMUNA = as.character(CODIGO_COMUNA)
  )


Adm2025_na_grouped <- Adm2025_na_grouped %>% mutate(year = 2025)
Adm2024_na_grouped <- Adm2024_na_grouped %>% mutate(year = 2024)
Adm2023_na_grouped <- Adm2023_na_grouped %>% mutate(year = 2023)
Adm2022_na_grouped <- Adm2022_na_grouped %>% mutate(year = 2022)
Adm2021_na_grouped <- Adm2021_na_grouped %>% mutate(year = 2021)

Adm_combined_na_2021_2025 <- bind_rows(
  Adm2025_na_grouped,
  Adm2024_na_grouped, 
  Adm2023_na_grouped, 
  Adm2022_na_grouped, 
  Adm2021_na_grouped
)
Adm_combined_na_21_25 <- Adm_combined_na_2021_2025 %>% 
  mutate(RBD = trimws(RBD)) %>%        # remove leading/trailing spaces
  filter(!is.na(RBD), RBD != "")


Adm_combined_na_21_25 <- Adm_combined_na_21_25 %>%
  select(
    year, everything()) %>%
  arrange(RBD, year)
View(Adm_combined_na_21_25)
```


## Save files

FILES 2021-2025 WITH 2012-2020 are really similar but i decided to separate it because have differetns RBD (idk why)

```{r}
write_csv(Adm_combined_21_25, "Adm_combined_21_25.csv")
write_csv(Adm_combined_na_21_25, "Adm_combined_na_21_25.csv")
```






